Forecasting the 2024 U.S. Presidential Election through Poll Aggregation and Adjustments for Poll Quality

### 2024-11-04 paper/paper.qmd
fix table and figure headings

## Basic SLR Model
```{r}
#| label: fig-model-pct-pollscore
#| fig-cap: "Comparison of Actual vs Predicted Percentage Support for Kamala Harris (Single Predictor: Pollscore)"
#| echo: false
#| warning: false

@@

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()
```

@@

```{r}
#| label: fig-model-pct-sample_size
#| fig-cap: "Actual vs. Predicted Values for pct Using sample_size as Predictor"
#| echo: false
#| warning: false
lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
predictions = predict(lm_model1, just_harris_data)
# summary(lm_model1)

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()

@@

In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample_size is the only predictor. This indicates that sample_size has limited impact on pct.

@@

```{r}
#| label: fig-model-MLR
#| fig-cap: "Actual vs Predicted Percentage Support for Kamala Harris (Multiple Predictors)"
#| echo: false
#| warning: false
ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()
```

@@

@fig-model-MLR compares actual pct values with the predicted pct values from the MLR model. Blue points represent individual comparisons between actual and predicted values. The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.

The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in pct. Most of the data points align well with the red dashed line, particularly within the middle range of pct values (between 40 and 60). This alignment indicates that the model performs well in this range, with predicted values correlating strongly with the actual observed outcomes.

@@

```{r}
#| label: fig-bayesian-model-actual-predict
#| fig-cap: "Predicted vs Actual Percentage Support (pct) for Kamala Harris in the Bayesian Model"
#| echo: false
#| warning: false
plot(test_data$pct, colMeans(predictions), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT")
abline(0, 1, col = "red")  # Add a 45-degree reference line
```

@@

```{r}
#| label: fig-bayesian-model-residual
#| fig-cap: "Residual Plot for the Bayesian Hierarchical Model Predictions"
#| echo: false
#| warning: false
residuals <- test_data$pct - colMeans(predictions)
plot(colMeans(predictions), residuals, 
     xlab = "Predicted Values", 
     ylab = "Residuals", 
     main = "Residual Plot")
abline(h = 0, col = "red")
```

@@

## Harris Voter Prediction Model 
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
```

@@

kable(head(result_summary, 10))
```

```{r}
#| label: fig-bayesian-model-result-CI
#| fig-cap: "U.S. Election Prediction Results"
#| echo: false
#| warning: false
ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "U.S. Election Prediction Results",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

## Trump Voter Prediction Model

The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris. Likewise, the Trump dataset is also split into training and testing data. The model outputs are below.

@@

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model (All Trump Data)",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

kable(head(result_summary, 10))

@@

```{r}
#| label: fig-model-trumpmlr-res
#| fig-cap: "MLR Trump Model does not Appear to Overfit"
#| echo: false
#| warning: false
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
plot(test_data_trump$pct, colMeans(predictions2), 
     xlab = "Actual PCT Values", 
     ylab = "Predicted  PCT Values", 
     main = "Predicted PCT vs Actual PCT (Test Data)")
abline(0, 1, col = "red")  # Add a 45-degree reference line
```

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalize outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias. 
```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% confidence interval ranges from around 42.5% to 47.2%. This confidence interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

@@

#The Ideal Survey

@@

Each state will have its own poll, but the polling methodology and the survey given will remain the same. Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes. Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for. However, these factors are almost impossible to determine while sending the survey. Therefore, the research team decided to ask demographic questions directly inside the survey. After data collection, the team will weigh data based on demographics. For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.

>>>>

## Basic SLR Model
```{r}
#| label: fig-model-pct-pollscore
#| fig-cap: "Pollscore has limited predictive accuracy for Kamala's voter percentage."
#| echo: false
#| warning: false

@@

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Single Linear Model (Only Pollscore Predictor) For Harris",
       x = "Actual Vote Percentage",
       y = "Predicted Vote Percentage") +
  theme_minimal()
```

@@

```{r}
#| label: fig-model-pct-sample_size
#| fig-cap: "Sample size has Limited Impact on Kamala's Voter Percentage"
#| echo: false
#| warning: false
lm_model1 = lm(pct ~ sample_size, data = just_harris_data)
predictions = predict(lm_model1, just_harris_data)
# summary(lm_model1)

ggplot(just_harris_data, aes(x = pct, y = predictions)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Sample Size as Predictor Model for Harris",
       x = "Actual Vote Percentage",
       y = "Predicted Vote Percentage") +
  theme_minimal()

@@

In @fig-model-pct-sample_size, the distribution of the points suggests that the model might not capture much variability in pct when sample size is the only predictor. This indicates that sample size has limited impact on pct.

@@

```{r}
#| label: fig-model-MLR
#| fig-cap: "The MLR Model captured significant variance in Kamala Harris' voter percentage"
#| echo: false
#| warning: false
ggplot(just_harris_data, aes(x = pct, y = fitted_value)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Multiple Predictors Model For Kamala Harris",
       x = "Actual Voter Percentage",
       y = "Predicted Voter Percentage") +
  theme_minimal()
```

@@

@fig-model-MLR compares actual percentage values with the predicted pct values from the MLR model. Blue points represent individual comparisons between actual and predicted values. The red dashed line represents the line of perfect prediction, where actual values would equal predicted values.

The overall distribution of points suggests that the MLR model has captured a substantial portion of the variance in Kamala Harris' voter percentage. Most of the data points align well with the red dashed line, particularly within the middle range of pct values (between 40 and 60). This alignment indicates that the model performs well in this range, with predicted values correlating strongly with the actual observed outcomes.

@@

```{r}
#| label: fig-bayesian-model-actual-predict
#| fig-cap: "The MLR Model Accruatley Predicts the Testing Dataset"
#| echo: false
#| warning: false
plot(test_data$pct, colMeans(predictions), 
     xlab = "Actual Voter Percentage Values", 
     ylab = "Predicted Voter Percentage Values", 
     main = "Multiple Predictors Kamala Harris Bayesian Model")
abline(0, 1, col = "red")
```

@@

```{r}
#| label: fig-bayesian-model-residual
#| fig-cap: "MLR Residuals Appear to Show No Trends"
#| echo: false
#| warning: false
residuals <- test_data$pct - colMeans(predictions)
plot(colMeans(predictions), residuals, 
     xlab = "Predicted Values", 
     ylab = "Residuals", 
     main = "Residual for the Bayesian Hierarchical Model Predictions")
abline(h = 0, col = "red")
```

@@

## Harris Voter Prediction Model 
```{r}
#| label: fig-bayesian-model-CI
#| fig-cap: "Kamala Harris is Predicted to get Approximatley 52% of Popular Vote"
#| echo: false
#| warning: false
# Plot random effects
plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95, 
     xlab = "Kamala Harris Popular Vote Percentage",
     title = "Estimated Posterior Distribution for the Intercept Parameter in the Bayesian Model")
```

@@

kable(head(result_summary, 10), 
      col.names = c("Acutal Kamala Percentage", 
                    "Predicted Percentage", 
                    "Lower Bound Credible Interval",
                    "Upper bound Credible Interval"))
```

```{r}
#| label: fig-bayesian-model-result-CI
#| fig-cap: "The Model Shows a Strong Alignment Between Actual and Predicted Vote Percentages"
#| echo: false
#| warning: false
ggplot(result_summary, aes(x = Actual_PCT, y = Predicted_PCT)) +
  geom_point() +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "U.S. Election Prediction Results",
    x = "Actual Percentage",
    y = "Predicted Percentage"
  )
```

@@

# Trump Voter Prediction Model

The multiple linear regression model (MLR) for Donald Trump will use the same variables, formula, and Bayesian approach as the one for Harris. Likewise, the Trump dataset is also split into training and testing data. The model results are similar to that of Harris's MLR model.

@@

```{r}
#| label: fig-model-trumpmlr-res
#| fig-cap: "Bayesian Hierarchical MLR Model of Trump Voting Percentage Captures Variability Without Overfitting"
#| fig-subcap: ["Trump Predictions (Test Data)", "Residuals for Predictions"]
#| echo: false
#| warning: false
#| layout-ncol: 2

predictions2 = posterior_predict(bayesian_model_train_trump, newdata = test_data_trump)
plot(test_data_trump$pct, colMeans(predictions2), 
     xlab = "Actual Voter Percentage Values", 
     ylab = "Predicted Voter Percentage Values")
abline(0, 1, col = "red")

residuals <- test_data$pct - colMeans(predictions)
plot(colMeans(predictions), residuals, 
     xlab = "Predicted Values", 
     ylab = "Residuals")
abline(h = 0, col = "red")
```

Based on @fig-model-trumpmlr-res, the test data predictions are also close to the actual values. This suggests that model can generalize outside data. Similarly, the distance between the prediction and the actual values do not appear to follow a pattern, suggesting that the error is due to randomness and not model bias.

# Trump Model Results

```{r}
#| label: fig-trumpmodelci
#| fig-cap: Donald Trump Expected to Recieve Approximatley 45% of the Vote
#| echo: false
#| warning: false
plot(bayesian_model_train_trump, pars = "(Intercept)", prob = 0.95)
```

Based on @fig-trumpmodelci, the model expects Trump to win slight less than 45% of the popular vote and the 95% credible interval ranges from around 42.5% to 47.2%. This credible interval is slightly larger than the model for Harris, implying that the Trump polling data might be less reliable.

```{r}
#| label: fig-model-trumpmlr
#| fig-cap: "MLR Trump Model Accounts For A Large Amount of Variability in Voter Percentage"
#| echo: false
#| warning: false
just_trump <- cleaned_data %>% filter(candidate_name == "Donald Trump")
predictions2 = posterior_predict(bayesian_model_train_trump, newdata = just_trump)
predicted_means = colMeans(predictions2)
predicted_intervals = apply(predictions2, 2, quantile, probs = c(0.025, 0.975))
result_summary = data.frame(
  Actual_PCT = just_trump$pct,
  Predicted_PCT = predicted_means,
  Lower_CI = predicted_intervals[1, ],
  Upper_CI = predicted_intervals[2, ]
)
ggplot(result_summary, aes(x = result_summary$Actual_PCT, y = result_summary$Predicted_PCT)) +
  geom_point(alpha = 0.7) +
  geom_errorbar(aes(ymin = Lower_CI, ymax = Upper_CI), width = 0.2, color = "blue") +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Actual vs Predicted Percentage For Trump MLR Model",
    x = "Actual PCT",
    y = "Predicted PCT"
  )
```

@@

kable(head(result_summary, 10), col.names = c("Acutal Harris Percentage", 
                    "Predicted Percentage", 
                    "Lower Bound Credible Interval",
                    "Upper bound Credible Interval"))

@@

@@

Based on the figures in @sec-data, the team identified three sources of bias beyond sample size and state: methodology, the timing of polls relative to the election date, and poll scores. From the analysis in Figure @fig-avg-pct-methodology, we found that polls employing a limited number of methods to reach and communicate with voters exhibited greater variability. This finding is plausible, as different demographic groups often prefer distinct communication methods; for example, older voters may favor phone surveys, while younger voters might prefer online surveys. Additionally, utilizing multiple communication channels provides pollsters with more opportunities to contact potential voters, thereby reducing non-response bias. Consequently, it is essential to incorporate methodology as a variable in the model, and we recommend that pollsters adopt diverse communication methods whenever possible.

@@

# The Ideal Survey

@@

Each state will have its own poll, but the polling methodology and the survey given will remain the same. Specifically, the polls will use stratified random sampling to make sure that participants reflect each state's counties in proportion to their population sizes. County proportion is used as the stratification variable because it is easily accessible. Additionally, counties possess inherent demographic and cultural characteristics that can influence their alignment with particular political parties. Ideally, stratification would also ensure race, gender, age, and other socioeconomic factors are also accounted for. However, these factors are almost impossible to determine while sending the survey. Therefore, the research team decided to ask demographic questions directly inside the survey. After data collection, the team will weigh data based on demographics. For example, if a certain county has a 30% black population (this statistic will be determined from the US census), but only 15% of the survey participants are black, then the pollsters may decide to count each black participant's responses twice.
